Periodic Traveling Waves in an Integro-Difference Equation With a Nonmonotone Growth Function and Strong Allee Effect

### 2021-06-21 paper_draft3.tex
more work

@@

@@

\begin{thm}  \label{theorem2}
If $\sup_{x\in\mathbb R}w_2(x)\leq \b$, then for all $\epsilon>0$, the sequence \ref{ptw} satisfies
\begin{equation}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n}(x)=1
\end{equation}
and
\begin{equation}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n+1}(x)=m
\end{equation}
and 
\begin{equation}
\lim_{n\to\infty}\sup_{x>n(c^*+\epsilon)}u_n(x)=0.
\end{equation}
\end{thm}

@@

\begin{equation} \begin{aligned}
\lim_{n\to\infty} \sup_{x>n(c^*+\epsilon)} u_n(x) 
&\leq \lim_{n\to\infty} \sup_{x>n(c^*+\epsilon)} \max\{w_1(x-nc^*),w_2(x-nc^*)\} \\
&= \lim_{n\to\infty} \sup_{x>n\epsilon} \max\{w_1(x),w_2(x)\} \\
&= \limsup_{x\to\infty} \max\{w_1(x),w_2(x)\} \\ 
&= \max\{\limsup_{x\to\infty} w_1(x),\limsup_{x\to\infty} w_2(x)\} \\
&= 0.
\end{aligned} \end{equation}
In the last calculation, the left hand side is known to be non-negative; therefore the limit is exactly equal to zero.
\end{proof}

@@

The reader can easily verify that the Laplace kernel satisfies hypotheses (H1) - (H3). The proof that is also satisfies (H4) is left in the appendix. The periodic traveling waves are given by

\begin{equation}
w_1(x) =   \begin{cases} 
1 - \frac{1}{2}e^{x} & x \leq 0 \\
\frac{1}{2}e^{-x} & x > 0
\end{cases}
\end{equation}
and
%\begin{equation}\label{laplacew2}
%w_2(x) = \begin{cases}
%m - \frac{1}{2} e^{x - \alpha} + \frac{1-\m}{2} e^{x - \beta} & x < \beta \\
%1 - \frac{1}{2} e^{x-\alpha} - \frac{1-\m}{2} e^{-(x-\beta)} & \beta \leq x \leq \alpha \\
%\frac{1}{2} e^{-(x-\alpha)} - \frac{1-\m}{2} e^{-(x-\beta)} & x > \alpha
%\end{cases}
%\end{equation}
\begin{equation} \label{w2laplace}
w_2(x) = \begin{cases}
m + C_1 e^x & x < x_\b \\
1 - C_2 e^x - C_3 e^{-x} & x_\b \leq x \leq x_\a \\
C_4 e^{-x} & x_\a < x 
\end{cases} \end{equation}
where
\begin{equation}
x_\a = \begin{cases} -\ln(2\a) & \a\leq \frac{1}{2} \\ \ln(2-2\a) & \a > \frac{1}{2} \end{cases} \quad
x_\b = \begin{cases} -\ln(2\b) & \b\leq \frac{1}{2} \\ \ln(2-2\b) & \b > \frac{1}{2} \end{cases} 
\end{equation}
%$x_\a=w_1^{-1}(\a)$, $x_\b=w_1^{-1}(\b)$, with
%$$ w_1^{-1}(p) = \begin{cases} -\ln(2p) & p\leq \frac{1}{2} \\ \ln(2-2p) & p > \frac{1}{2} \end{cases} $$
The constants $C_1,C_2,C_3$, and $C_4$ are continuous functions of the growth parameters $\alpha$, $\beta$, and $\mu$, with $C_2,C_3,C_4\geq0$. They can be expressed as piecewise expressions:
\newcommand{\Conecaseone}{\b(1-\m)-\a}
\newcommand{\Conecasetwo}{\frac{1-\m-4\a(1-\b)}{4(1-\b)}}
\newcommand{\Conecasethree}{- \frac{1-\b+\m(1-\a)}{4(1-\a)(1-\b)} }

>>>>

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{figures/fig2A.png}
  \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{figures/fig2B.png}
  \label{fig:sub2}
\end{subfigure}
\caption{Plots of $w_1(x)$ (left) and $w_2(x)$ (right) with the Laplace dispersal kernel and growth parameters $\alpha=0.3$, $\mu=0.5$, and $\beta=0.8$}
\label{fig:test}
\end{figure}

@@

The condition that $\sup_{x\in\mathbb R}w_2(x)\leq \b$ may be difficult to verify for some choices of dispersal kernel. Thus, we provide a sufficient condition for this assumption:
\begin{lem} \label{lemma}
Let $x_\a$ and $x_\b$ be as defined in Lemma \ref{lemma1}. If $k(x)$ is unimodal and
$$ \m +  (1-\mu)\int_{-(x_\a-x_\b)/2}^{(x_\a-x_\b)/2} k (y) \, dy \leq \b $$
then $\sup_{x\in\mathbb R}w_2(x) \leq \b$.
\end{lem}

\begin{proof}
For all $x\in\mathbb R$, we have
\begin{equation} \begin{aligned}
w_2(x) &= \int_{-\infty}^{\infty} k(x-y) g(w_1(y)) \, dy  \\
&\leq\int_{-\infty}^{\infty} k(x-y) (\m + (\m-1) \mathbf 1_{[x_\b,x_\a]}(y) )\, dy \\
&= \m + \int_{-\infty}^{\infty} k(x-y) (\m-1) \mathbf 1_{[x_\b,x_\a]}(y) \, dy \\
&= \m +  (1-\mu)\int_{x_\b}^{x_\a} k (x-y) \, dy .
\end{aligned}
\end{equation}

Taking the maximum over all $x$, we may shift the function horizontally so that the bounds of integration are symmetric. Since $k(x)$ is unimodal, the integral is then maximized for $x=0$.
\begin{equation} \begin{aligned}
\sup_{x\in\mathbb R} w_2(x) &\leq \sup_{x\in\mathbb R} \left( \m +  (1-\mu)\int_{x_\b}^{x_\a} k (x-y) \, dy \right) \\
&= \sup_{x\in\mathbb R} \left( \m +  (1-\mu)\int_{-(x_\a-x_\b)/2}^{(x_\a-x_\b)/2} k (x-y) \, dy \right) \\
&= \m +  (1-\mu)\int_{-(x_\a-x_\b)/2}^{(x_\a-x_\b)/2} k (y) \, dy.
\end{aligned}
\end{equation}
\end{proof}

@@

\begin{figure}[h!] 
\centering
  \caption{Numerical simulation of the integro-difference equation with Laplace dispersal kernel, $\alpha=0.2$, $\mu=0.3$, and $\beta=0.8$. Later timepoints are colored in red, with earlier timepoints colored in green.}
\label{fig3}
  \includegraphics{figures/fig3.png}
\end{figure}

Figure \ref{fig3} shows the propagation of the periodic traveling wave solution for a particular choice of dispersal kernel and growth parameters.

\begin{thm}  \label{theorem2}
If $\sup_{x\in\mathbb R}w_2(x)\leq \b$, then for all $\epsilon>0$, the sequence \ref{ptw} satisfies
\begin{equation}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n}(x)=1
\end{equation}
and
\begin{equation}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n+1}(x)=\m
\end{equation}
and 
\begin{equation}
\lim_{n\to\infty}\sup_{x>n(c^*+\epsilon)}u_n(x)=0.
\end{equation}
\end{thm}

@@

\begin{equation} \begin{aligned}
\lim_{n\to\infty} \sup_{x>n(c^*+\epsilon)} u_n(x) 
&\leq \lim_{n\to\infty} \sup_{x>n(c^*+\epsilon)} \sup\{w_1(x-nc^*),w_2(x-nc^*)\} \\
&= \lim_{n\to\infty} \sup_{x>n\epsilon} \sup\{w_1(x),w_2(x)\} \\
&= \limsup_{x\to\infty} \sup\{w_1(x),w_2(x)\} \\ 
&= \sup\{\limsup_{x\to\infty} w_1(x),\limsup_{x\to\infty} w_2(x)\} \\
&= 0.
\end{aligned} \end{equation}
In the last calculation, the left hand side is known to be non-negative; therefore the limit is exactly equal to zero.
\end{proof}

@@

The reader can easily verify that the Laplace kernel satisfies hypotheses (H1) - (H3); the proof of (H4) is left in the appendix.

The periodic traveling wave profiles are given by

\begin{equation}
w_1(x) =   \begin{cases} 
1 - \frac{1}{2}e^{x} & x \leq 0 \\
\frac{1}{2}e^{-x} & x > 0
\end{cases}
\end{equation}
and
%\begin{equation}\label{laplacew2}
%w_2(x) = \begin{cases}
%m - \frac{1}{2} e^{x - \alpha} + \frac{1-\m}{2} e^{x - \beta} & x < \beta \\
%1 - \frac{1}{2} e^{x-\alpha} - \frac{1-\m}{2} e^{-(x-\beta)} & \beta \leq x \leq \alpha \\
%\frac{1}{2} e^{-(x-\alpha)} - \frac{1-\m}{2} e^{-(x-\beta)} & x > \alpha
%\end{cases}
%\end{equation}
%\begin{equation} \label{w2laplace}
%w_2(x) = \begin{cases}
%m + C_1 e^x & x < x_\b \\
%1 - C_2 e^x - C_3 e^{-x} & x_\b \leq x \leq x_\a \\
%C_4 e^{-x} & x_\a < x 
%\end{cases} \end{equation}
%or equivalently

\begin{equation} \label{w2laplace}
w_2(x) = \begin{cases}
m + \left( \frac{1-\m}{2}e^{-x_\b} - \frac{1}{2}e^{-x_\a} \right) e^x & x < x_\b \\
1 - \left( \frac{1}{2} e^{-x_\a} \right) e^x - \left( \frac{1-\m}{2} e^{x_\b} \right)  e^{-x} & x_\b \leq x \leq x_\a \\
\left( \frac{1}{2} e^{x_\a} - \frac{1-\m}{2} e^{x_\b} \right) e^{-x} & x_\a < x 
\end{cases} \end{equation}

where
\begin{equation}
x_\a = \begin{cases} -\ln(2\a) & \a\leq \frac{1}{2} \\ \ln(2-2\a) & \a > \frac{1}{2} \end{cases}, \quad
x_\b = \begin{cases} -\ln(2\b) & \b\leq \frac{1}{2} \\ \ln(2-2\b) & \b > \frac{1}{2} \end{cases} 
\end{equation}
This leads to an equation for the mean wave speed
\begin{equation}
c^* = \begin{cases}
\frac{1}{2}x_\a  + \frac{1}{2}\log \left( \frac{1 - (1-\m) e^{x_\b-x_\a}}{2\a} \right) & \a \leq w_2(x_\a) \\
\frac{1}{2}x_\a + \frac{1}{2} \log \left( 1-\a + \sqrt{(1-\a)^2 - (1-\mu) e^{x_\b-x_\a}} \right) & w_2(x_\a) < \a < w_2(x_\b) \\
\frac{1}{2}x_\a + \frac{1}{2} \log \left( \frac{2(\a-\m)}{(1-\m)e^{x_\a-x_\b}-1} \right) & \a \geq w_2(x_\b)
\end{cases} 
\end{equation}
and

\begin{equation} \begin{aligned}
C_1 &= \frac{1-\m}{2}e^{-x_\b} - \frac{1}{2}e^{-x_\a} \\
C_2 &= \frac{1}{2} e^{-x_\a} \\
C_3 &= \frac{1-\m}{2} e^{x_\b} \\
C_4 &= \frac{1}{2} e^{x_\a} - \frac{1-\m}{2} e^{x_\b}
\end{aligned} \end{equation}

%$x_\a=w_1^{-1}(\a)$, $x_\b=w_1^{-1}(\b)$, with
%$$ w_1^{-1}(p) = \begin{cases} -\ln(2p) & p\leq \frac{1}{2} \\ \ln(2-2p) & p > \frac{1}{2} \end{cases} $$
The constants $C_1,C_2,C_3$, and $C_4$ are continuous functions of the growth parameters $\alpha$, $\beta$, and $\mu$, with $C_2,C_3,C_4\geq0$. They can be expressed as piecewise expressions:
\newcommand{\Conecaseone}{\b(1-\m)-\a}
\newcommand{\Conecasetwo}{\frac{1-\m-4\a(1-\b)}{4(1-\b)}}
\newcommand{\Conecasethree}{- \frac{1-\b+\m(1-\a)}{4(1-\a)(1-\b)} }
